AtmoSwing (v1.4): Analog Technique Model for Statistical weather forecastING

### 2015-10-21 AtmoSwing - Geoscientific Model Development/template.tex
Merging the evolution of the method with references.

\subsection{Evolution of the method}
\label{sec:method:evolution}

The most common usage of the analogue method is for precipitation forecasting. Thus, the evolution presented hereafter focuses on this predictand.

Taking advantage of higher-performance IT tools \citet{Guilbaud1997} stopped using PCA in order to work directly with the raw data interpolated on grids, which was an improvement, especially when considering the S1 criterion (already used for analogues selection by\citealp{Wilson1980} and \citealp{Woodcock1980}) instead of the Euclidean distance. This score gives importance to the similarity of atmospheric circulation rather than absolute values of the pressure field. After comparing this criterion to other criteria from the literature, or combinations of these, \citet{Guilbaud1998} found that S1 is the most efficient. They also pointed out the benefit of using 2 atmospheric levels instead of one (also instead of the thickness), and to also consider 2 temporal windows \citep{Obled2002}. When using multiple predictors, they combined the criteria values calculated on each atmospheric level and each temporal window using an arithmetic mean. Finally, they also suggested using a second level of analogy on lower layers moisture and introduced the Ranked Probability Score \citep{Epstein1969} to assess the forecast quality, which was found to be more relevant that the Brier score \citep{Brier1950}.

\citet{Bontron2004} could benefit from the newly available NCEP/NCAR reanalysis data. This dataset being more homogeneous, much longer, and contained more variables than what was previously used, he could systematically assess a large selection of variables in the analogue method combined with different criteria. He thus confirmed the interest of the geopotential height along with the S1 criterion for the first level of analogy. He has also shown that the choice of the temporal window (time of observation) has a higher importance than the atmospheric level for the performance of the method. By assessing multiple combinations of atmospheric levels and temporal windows, he concluded that the coupled geopotential heights at 1000~hPa \& 12~h and 500~hPa \& 24~h provided the best performance for the studied region.

He also assessed several variables for the second level of analogy and identified the moisture variables (precipitable water and relative humidity) and vertical velocity to be the most relevant. Finally, he found that the product of relative humidity and precipitable water gave the best performance.

\citet{BenDaoud2010} applied the analogue method in the context of large flatter catchments, out of an Alpine environment (Sa\^{o}ne, Seine). He evaluated several variables usually used in weather forecasting, amongst them the air temperature and the vertical velocity that presented a potential interest. The temperature (at the nearest grid point, on the 925~hPa (36~h) and 600~hPa (12~h) levels) could replace the rigid selection of $\pm$ 60 days around the target date by a more dynamic screening of similar situations in terms of air masses. The seasonal effect is indeed also present in the temperature data. The vertical velocity was used as an additional level of analogy, between the analogy of atmospheric circulation and the one on moisture, to account for fronts. Finally, he selected the moisture variables at 700~hPa instead of 850~Pa.

% TODO : Sabine
% TODO : Jérémie Chardon
% TODO : Horton

@@

\subsection{References for precipitation forecasting}

The evolution of the analogue method for precipitation forecasting led to several improvements. However, some of these are specific for a certain region and are not relevant for others, and some perform better depending on the lead time. Thereby, we cannot provide a single best parameterization for the method, neither rank them from worst to best. We will thus provide the references implementations of the most recent method variant, in the chronological order they appeared.

The first parameterization comes from the work of \citet{Guilbaud1997} and \citet{Obled2002}. It consists of a single level of analogy on the atmospheric circulation. The score S1 (Teweles-Wobus, Eq. \ref{eq:S1}) is processed on the geopotential height at 1000~hPa (Z1000) and 700~hPa (Z700) or 500~hPa (Z500) (Table \ref{table:R0}), both at 00~h and 24~h. The number of analogues considered is 50.

@@

The analogy on the atmospheric circulation (Table \ref{table:R1}) proposed by \citet{Bontron2004} is still, at the time of writing, often used, and considered as a reference for benchmarking new parameterizations. He noticed that the choice of the temporal window was more important than the atmospheric level. The method consists of the geopotential height on two atmospheric levels : 1000~hPa (Z1000) at 12~h and 500~hPa (Z700) at 24~h.

@@

\citep{Bontron2004} then added a second level of analogy (Table \ref{table:R2}) on a moisture index : the product of relative humidity at 850~hPa (RH850) and total precipitable water (TPW). This index does not represent an actual physical quantity, but expresses the water content of the column and its proximity to saturation.

@@

The next reference method was developed by \citet{BenDaoud2010} and introduces a change in the preselection. The calendar selection ($\pm 60$ days around the target date) was then replaced by a preselection based on air temperature at 925~hPa (36~h) and 600~hPa (12h). The number of preselected dates is equivalent to the number of days we would have chosen with the calendar approach, and thus depends on the archive size: $N_{0} = 60 \cdot 2 \cdot n_{a}$ where $n_{a}$ is the number of years in the calibration period. This analogy on air temperature is performed at the closest point to the catchment

\citet{BenDaoud2010} also reconsidered the parameters of the moisture index and ended up with both 925~hPa and 700~hPa levels, at 12~h and 24~h, or at every time step (6~h) between 06~h and 30~h (Table \ref{table:R4b}).

>>>>

@@

\subsection{Evolution of the method and examples}
\label{sec:method:evolution}

The evolution of the analogue method for precipitation forecasting led to several improvements. However, some of these are specific for a certain region and are not relevant for others, and some perform better depending on the lead time. Thereby, we cannot provide a single best parameterization for the method, neither rank them from worst to best. We will thus provide some examples of parameterizations of the most recent method variant, in the chronological order they appeared. The most common usage of the analogue method is for precipitation forecasting. Thus, the evolution presented hereafter focuses on this predictand.

Taking advantage of higher-performance IT tools \citet{Guilbaud1997} stopped using PCA in order to work directly with the raw data interpolated on grids, which was an improvement, especially when considering the S1 criterion (already used for analogues selection by\citealp{Wilson1980} and \citealp{Woodcock1980}) instead of the Euclidean distance. This score gives importance to the similarity of atmospheric circulation rather than absolute values of the pressure field. After comparing this criterion to other criteria from the literature, or combinations of these, \citet{Guilbaud1998} found that S1 is the most efficient. They also pointed out the benefit of using 2 atmospheric levels, 1000~hPa (Z1000) and 700~hPa (Z700) or 500~hPa (Z500), instead of one (also instead of the thickness), and to also consider 2 temporal windows, 00~h and 24~h \citep{Obled2002}. The number of analogues considered is 50 (Table \ref{table:R0}). When using multiple predictors, they combined the criteria values calculated on each atmospheric level and each temporal window using an arithmetic mean. Finally, they also suggested using a second level of analogy on lower layers moisture and introduced the Ranked Probability Score \citep{Epstein1969} to assess the forecast quality, which was found to be more relevant that the Brier score \citep{Brier1950}.

@@

\citet{Bontron2004} could benefit from the newly available NCEP/NCAR reanalysis data. This dataset being more homogeneous, much longer, and contained more variables than what was previously used, he could systematically assess a large selection of variables in the analogue method combined with different criteria. He thus confirmed the interest of the geopotential height along with the S1 criterion for the first level of analogy. He has also shown that the choice of the temporal window (time of observation) has a higher importance than the atmospheric level for the performance of the method. By assessing multiple combinations of atmospheric levels and temporal windows, he concluded that the coupled geopotential heights at 1000~hPa (Z1000) \& 12~h and 500~hPa (Z500) \& 24~h provided the best performance for the studied region (Table \ref{table:R1}). The analogy on the atmospheric circulation proposed by \citet{Bontron2004} is still, at the time of writing, often used operationally, and considered as a reference for benchmarking new parameterizations \citep{Horton2012, Daoud2015}.

@@

\citep{Bontron2004} also assessed several variables for the second level of analogy and identified the moisture variables (precipitable water and relative humidity) and vertical velocity to be the most relevant. Finally, he found that a moisture index made of the product of relative humidity at 850~hPa (RH850) and total precipitable water (TPW) gave the best performance (Table \ref{table:R2}). This index does not represent an actual physical quantity, but expresses the water content of the column and its proximity to saturation.

@@

\citet{BenDaoud2010} applied the analogue method in the context of large floodplains, out of an Alpine environment (Sa\^{o}ne, Seine). He evaluated several variables usually used in weather forecasting, amongst them the air temperature and the vertical velocity that presented a potential interest. The temperature (at the nearest grid point, on the 925~hPa (36~h) and 600~hPa (12~h) levels) could replace the rigid calendar preselection of $\pm$ 60 days around the target date by a more dynamic screening of similar situations in terms of air masses. The seasonal effect is indeed also present in the temperature data. The number of preselected dates is equivalent to the number of days we would have chosen with the calendar approach, and thus depends on the archive size: $N_{0} = 60 \cdot 2 \cdot n_{a}$ where $n_{a}$ is the number of years in the calibration period. \citet{BenDaoud2010} also reconsidered the parameters of the moisture index and ended up with both 925~hPa and 700~hPa levels, at 12~h and 24~h, or at every time step (6~h) between 06~h and 30~h (Table \ref{table:R4b}).
